\thetitle

### 2020-11-09 sun_phases/spin_qudit_tomo.tex
minor cleaning up of some text

In this work, we provide an efficient framework for performing spin qudit tomography using only measurements of spin projection onto different spatial axes.
We first introduce, in Section \ref{sec:transition_ops}, a basis of operators that have a simple physical interpretation, and transform nicely under rotations in 3D space, thereby respecting the symmetries of the tomography task at hand.
We then map this quantum tomography problem onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and mathematical machinery that has been developed for this task \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
In particular, this mapping allows us to bound the minimum number of measurement axes necessary to perform full tomography on a $d$-level spin qudit, $r_d$, by $2d-1\le r_d\le d^2$.

@@

Consider a $d$-state spin qudit with total spin $s\equiv\frac{d-1}{2}$.
The defining property of a spin qudit is the fact that it describes an angular momentum degree of freedom, which has specific implications for how a spin qudit should transform under the group SO(3) of rotations in 3D space.
Due to the central importance of these transformation rules for a spin qudit, we first identify a basis of operators that transforms nicely under 3D rotations.
One such basis is provided by the {\it transition operators} (also known as polarization operators \cite{kryszewski2006positivity, bertlmann2008bloch}), defined by
\begin{align}
  T_{\ell m} \equiv \sqrt{\f{2\ell+1}{2s+1}} \sum_{\mu,\nu=-s}^s
  \bk{s\mu;\ell m|s\nu} \op{\nu}{\mu},
  \label{eq:trans_op}
\end{align}
where $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,d-1}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$, such that there are $d^2$ operators in total.
For brevity, we will generally treat the value of $d$ (or $s$) as constant but arbitrary throughout this work, and suppress any explicit dependence of quantities or operators such as $T_{\ell m}$ on $d$.
The transition operators are orthonormal with respect to the trace inner product, and transform nicely under conjugation:
\begin{align}
  \obk{T_{\ell m }|T_{\ell'm'}}
  = \delta_{\ell\ell'} \delta_{mm'},
  &&
  T_{\ell m}^\dag = \p{-1}^m T_{\ell,-m},
\end{align}
where for any $d\times d$ matrix $X=\sum_{\mu,\nu} X_{\mu\nu} \op{\mu}{\nu}$ we define the $d^2$-component vector $\oket{X} \equiv \sum_{\mu,\nu} X_{\mu\nu} \ket{\mu\nu}$; $\obra{X}$ is the conjugate transpose of $\oket{X}$, such that $\obk{X|Y}=\tr\p{X^\dag Y}$; and $\delta_{kk'}\equiv 1$ if $k=k'$ and $0$ otherwise.
As a consequence, we can expand any density operator $\rho$ in the transition operator basis as
\begin{align}
  \rho = \sum_{\ell=0}^{2s} \sum_{m=-\ell}^\ell \rho_{\ell m} T_{\ell m},
  &&
  \rho_{\ell m} \equiv \obk{T_{\ell m}|\rho},
  \label{eq:trans_state}
\end{align}
where $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
The transition operators can be physically interpreted as the transition $\ket\psi\to T_{\ell m}\ket\psi$ induced on the state $\ket\psi$ of a spin-$s$ system by the absorption of a spin-$\ell$ particle with spin projection $m$ onto a fixed quantization axis.
In analogy with the spherical harmonics $Y_{\ell m}$, we will refer to the index $\ell$ as the {\it degree} and $m$ as the {\it order} of $T_{\ell m}$.

@@

If the function $f$ is {\it band-limited} at degree $L$, which is to say that $f_{\ell m}=0$ for all $\ell\ge L$, then the full recovery problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points \cite{freeden2008spherical, freeden2018spherical}.
The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states are essentially ``band-limited'' at degree $d$ \footnote{Spin qudit states can be faithfully represented by quasi-probability distributions on a sphere.  For a $d$-level spin, these quasi-probability distributions are band-limited at degree $d$.}.
Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f = \sum_m f_m Y_{\ell m}$, which is possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are transition operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error scale $\epsilon_V^2$ defined in Eq.~\eqref{eq:scale}.
If $\abs{V}$ is too large for such optimization  to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest error scale $\epsilon_V^2$.
After choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{\Pi_{\v v\mu}}_\rho$ by making measuring spin projection along every axis $\v v\in V$.
These spin projections determine the expectation values $\bk{T_{\v v\ell,0}}_\rho$, whose estimates can be used to construct the state $\tilde\rho$ defined in Eq.~\eqref{eq:reconstructed_state}.
If spin projection was measured $n$ times along every axis, then the reconstructed state $\tilde\rho$ approximates the true state $\rho$ to within a mean squared (Euclidean) distance $\epsilon_V^2/n$.
Furthermore, if $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

@@

To this end, we fix a particular set of measurement axes $V$, and consider performing $n$ measurements of spin projection along every axis $\v v\in V$, for a total of $N=\abs{V}\times n$ measurements.
Such a procedure is equivalent to making $N$ local measurements of the $N$-fold product state $\rho^{\otimes N}$.
For convenience, we index the tensor factors of $\rho^{\otimes N}$ by the integers $\p{i,j}$, with $i\in\set{1,2,\cdots,\abs{V}}$ specifying a measurement axis $\v v_i\in V$ and $j\in\set{1,2,\cdots,n}$ specifying the copy of $\rho$ prepared for the $j$-th measurement spin projection along a particular axis.
We then define the projectors $\Pi_{i\mu}\equiv\op{\mu_{\v v_i}}$ onto states $\ket{\mu_{\v v_i}}$ with definite spin projection $\mu$ onto axis $\v v_i\in V$, and define $\Pi_{i\mu}^j$ to be an $N$-qudit operator with $\Pi_{i\mu}$ on the $\p{i,j}$-th tensor factor and the identity elsewhere.
We denote the experimental outcome of measuring $\Pi_{i\mu}$ in the $\p{i,j}$-th copy of $\rho$ by $\tilde \Pi_{i\mu}^j\in\set{0,1}$.
In other words, $\tilde \Pi_{i\mu}^j$ is the ``single-shot estimate'' of $\Pi_{i\mu}$, with $\tilde \Pi_{i\mu}^j=1$ if outcome $\mu$ was observed on the $\p{i,j}$-th experimental trial, and $\tilde \Pi_{i\mu}^j=0$ otherwise.
An empirical estimate of the expectation value $\bk{\Pi_{i\mu}}_\rho$ is then provided by the fraction of times that outcome $\mu$ was observed when measuring spin projection along axis $\v v_i$, that is
\begin{align}
  \tilde \Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \tilde \Pi_{i\mu}^j
  \approx \f1n \sum_{j=1}^n \tr\p{\Pi_{i\mu}^j \rho^{\otimes N}}
  = \tr\p{\Pi_{i\mu} \rho}.
\end{align}
For reasons that will become clear shortly, it will be useful to think of $\tilde \Pi_{i\mu}$ as an empirical estimate of $\bk{\bar \Pi_{i\mu}}_{\rho^{\otimes N}}$, where
\begin{align}
  \bar \Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \Pi_{i\mu}^j
  \label{eq:mean_proj}
\end{align}
is the average of $\Pi_{i\mu}$ applied to all copies of $\rho$ for which spin projection is measured along the axis $\v v_i$.

@@

Due to finite sampling error (i.e.~shot noise), an empirical estimate $\tilde\O$ of the observable $\O$ will have statistical error with mean zero:
\begin{align}
  \epsilon_\O \equiv \tilde\O - \bk{\O},
\end{align}
where the single brackets $\bk{\cdot}$ denote an expectation value with respect to the measured quantum state.
On average, this statistical error will be zero, which is to say that
\begin{align}
  \bbk{\epsilon_\O} = \Bbk{\tilde\O - \bk{\O}}
  = \bk{\O - \bk{\O}}
  = 0,
\end{align}
where the double brackets $\bbk{\cdot}$ to denote statistical averaging over experimental trials.
However, the covariance between statistical errors $\epsilon_\O$ and $\epsilon_\Q$ on the empirical estimates $\tilde\O$ and $\tilde\Q$ of observables $\O$ and $\Q$ is
\begin{align}
  \bbk{\epsilon_\O \epsilon_\Q}
  = \Bbk{\p{\tilde\O - \bk{\O}} \p{\tilde\Q - \bk{\Q}}}
  = \bk{\p{\O - \bk{\O}} \p{\Q - \bk{\Q}}}
  = \bk{\O\Q} - \bk{\O} \bk{\Q}.
\end{align}
In the context of spin qudit tomography, we can therefore define the statistical error
\begin{align}
  \epsilon_{i\mu}
  \equiv \tilde \Pi_{i\mu} - \bk{\Pi_{i\mu}}_{\rho}
  = \tilde \Pi_{i\mu} - \bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}
\end{align}
in the empirical estimate of $\bk{\Pi_{i\mu}}_{\rho}$ or $\bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}$, and use Eq.~\eqref{eq:mean_proj} to expand
\begin{align}
  \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
  = \bk{\bar\Pi_{i\mu} \bar\Pi_{i'\mu'}}_{\rho^{\otimes N}}
  - \bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}
  \bk{\bar\Pi_{i'\mu'}}_{\rho^{\otimes N}}
  = \f1{n^2} \sum_{j,j'=1}^n
  \sp{\bk{\Pi_{i\mu}^j \Pi_{i'\mu'}^{j'}}_{\rho^{\otimes N}}
    - \bk{\Pi_{i\mu}^j}_{\rho^{\otimes N}}
    \bk{\Pi_{i'\mu'}^{j'}}_{\rho^{\otimes N}}}.
  \label{eq:proj_cov_start}
\end{align}
If $\p{i,j}\ne\p{i',j'}$, then $\Pi_{i\mu}^j$ and $\Pi_{i'\mu'}^{j'}$ address different tensor factors of the product state $\rho^{\otimes N}$, so the expectation value of their product factorizes due to the identity $\tr\sp{\p{A\otimes B}\p{A'\otimes B'}} = \tr\p{AA'}\times\tr\p{BB'}$.
The terms in Eq.~\eqref{eq:proj_cov_start} with $\p{i,j}\ne\p{i',j'}$ therefore vanish, so
\begin{align}
  \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
  &= \delta_{ii'} \times \f1{n^2} \sum_{j=1}^n \sp{\bk{\Pi_{i\mu}^j \Pi_{i\mu'}^j}_{\rho^{\otimes N}}
    - \bk{\Pi_{i\mu}^j}_{\rho^{\otimes N}}
    \bk{\Pi_{i\mu'}^j}_{\rho^{\otimes N}}} \\
  &= \delta_{ii'} \times \f1n \sp{\bk{\Pi_{i\mu} \Pi_{i\mu'}}_\rho
    - \bk{\Pi_{i\mu}}_\rho \bk{\Pi_{i\mu'}}_\rho} \\
  &= \delta_{ii'} \times \f1n \sp{\delta_{\mu\mu'} p_\mu^i - p_\mu^i p_{\mu'}^i},
\end{align}
where $p_\mu^i \equiv \bk{\Pi_{i\mu}}_\rho = \bk{\mu_{\v v_i}|\rho|\mu_{\v v_i}}$ are the diagonal entries of $\rho$ in the basis of spin projections along axis $\v v_i$.

@@

Equipped with the bound on variances in Eq.~\eqref{eq:variance_bound}, we again consider the block-diagonalized form of the measurement matrix derived in Section \ref{sec:bound}:
\begin{align}
  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
  &&
  M_{V\ell} = \sum_{m,\v v} D^\ell_{m,0} \op{m}{\v v}.
\end{align}
The block diagonal structure of this matrix implies that, rather than an integer index $k\in\set{1,2,\cdots,d^2}$, we can identify the singular values and right singular vectors of $UM_V$ (and equivalently $M_V$) by a block index $\ell\in\set{0,1,\cdots,d-1}$ and a sub-index $m=\set{-\ell,-\ell+1,\cdots,\ell}$.
Making the replacement $k\to\p{\ell,m}$, we can thus repeat the derivations through Eq.~\eqref{eq:bound_eps} of Section \ref{sec:bound} without the assumption of Eq.~\eqref{eq:error_assumption}, finding that
\begin{align}
  \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
  = \sum_{i,i'} q_{\ell m i}^V \p{q_{\ell m i'}^V}^*
  \bbk{\epsilon_{i\ell} \epsilon_{i'\ell}}
  = \sum_i \abs{q_{\ell m i}^V}^2
  \bbk{\epsilon_{i\ell}^2}
  < \f{2\ell+1}{nd} \times \gamma_\ell^2
  \times \p{1-\delta_{\ell,0}},
  \label{eq:bound_single}
\end{align}
where $i$ indexes an axis $\v v_i\in V$.
Altogether, we thus find that $n$ measurements of spin projection along every axis $\v v\in V$ allows us to construct an empirical estimate $\tilde\rho_V$ of an unknown state $\rho$ that is accurate to within a mean squared distance
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \sum_{\ell,m} \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
  < \f1{nd} \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2.
  \label{eq:bound_full}
\end{align}
We note that whereas the bound on $\bbk{\epsilon_{i\ell}^2}$ in Eq.~\eqref{eq:variance_bound} is tight, the bounds in Eq.~\eqref{eq:bound_single} and Eq.~\eqref{eq:bound_full} are not: the upper bound on $\bbk{\epsilon_{i\ell}^2}$ cannot be achieved simultaneously for all axes $\v v_i$ or degrees $\ell$.
In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound_full}.

>>>>

In this work, we provide an efficient framework for performing spin qudit tomography using only measurements of spin projection onto different spatial axes.
We first introduce, in Section \ref{sec:transition_ops}, a basis of operators that have a simple physical interpretation, and transform nicely under rotations in 3D space, thereby respecting the symmetries of the tomography task at hand.
We then map this quantum tomography problem onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and mathematical machinery that has been developed for this task \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
In particular, this mapping allows us to bound the minimum number of measurement axes necessary to perform full tomography on a $d$-level spin qudit, $r_d^{\t{min}}$, by $2d-1\le r_d^{\t{min}}\le d^2$.

@@

Consider a $d$-state spin qudit with total spin $s\equiv\frac{d-1}{2}$.
The defining property of a spin qudit is the fact that it describes an angular momentum degree of freedom, which has specific implications for how a spin qudit should transform under the group SO(3) of rotations in 3D space.
Due to the central importance of these transformation rules for a spin qudit, we first identify a basis of operators that transforms nicely under 3D rotations.
One such basis is provided by the {\it transition operators} (also known as polarization operators \cite{kryszewski2006positivity, bertlmann2008bloch}), defined by
\begin{align}
  T_{\ell m} \equiv \sqrt{\f{2\ell+1}{2s+1}} \sum_{\mu,\nu=-s}^s
  \bk{s\mu;\ell m|s\nu} \op{\nu}{\mu},
  \label{eq:trans_op}
\end{align}
where $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,d-1}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$, such that there are $d^2$ operators in total.
For brevity, we will generally treat the value of $d$ (or $s$) as constant but arbitrary throughout this work, and suppress any explicit dependence of quantities or operators such as $T_{\ell m}$ on $d$.
The transition operators are orthonormal with respect to the trace inner product, and transform nicely under conjugation:
\begin{align}
  \obk{T_{\ell m }|T_{\ell'm'}}
  = \delta_{\ell\ell'} \delta_{mm'},
  &&
  T_{\ell m}^\dag = \p{-1}^m T_{\ell,-m},
\end{align}
where for any $d\times d$ matrix $X=\sum_{\mu,\nu} X_{\mu\nu} \op{\mu}{\nu}$ we define the $d^2$-component vector $\oket{X} \equiv \sum_{\mu,\nu} X_{\mu\nu} \ket{\mu\nu}$; $\obra{X}$ is the conjugate transpose of $\oket{X}$, such that $\obk{X|Y}=\tr\p{X^\dag Y}$; and $\delta_{kk'}\equiv 1$ if $k=k'$ and $0$ otherwise.
These properties of the transition operators allow us to expand any density operator $\rho$ in the transition operator basis as
\begin{align}
  \rho = \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell \rho_{\ell m} T_{\ell m},
  &&
  \rho_{\ell m} \equiv \obk{T_{\ell m}|\rho},
  \label{eq:trans_state}
\end{align}
where $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
The transition operators can be physically interpreted as the transition $\ket\psi\to T_{\ell m}\ket\psi$ induced on the state $\ket\psi$ of a spin-$s$ system by the absorption of a spin-$\ell$ particle with spin projection $m$ onto a fixed quantization axis.
In analogy with the spherical harmonics $Y_{\ell m}$, we will refer to the index $\ell$ as the {\it degree} and $m$ as the {\it order} of $T_{\ell m}$.

@@

If the function $f$ is {\it band-limited} at degree $L$, which is to say that $f_{\ell m}=0$ for all $\ell\ge L$, then the full recovery problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points \cite{freeden2008spherical, freeden2018spherical}.
The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states are essentially ``band-limited'' at degree $d$ \footnote{Spin qudit states can be faithfully represented by quasi-probability distributions on a sphere.  For a $d$-level spin, these quasi-probability distributions are band-limited at degree $d$.}.
Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f = \sum_m f_m Y_{\ell m}$, which provably is possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are transition operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error scale $\epsilon_V^2$ defined in Eq.~\eqref{eq:scale}.
If $\abs{V}$ is too large for such optimization  to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest error scale $\epsilon_V^2$.
After choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{\Pi_{\v v\mu}}_\rho$ by making measuring spin projection along every axis $\v v\in V$.
These spin projections determine the expectation values $\bk{T_{\v v\ell,0}}_\rho$, whose estimates can be used to construct the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state}.
If spin projection was measured $n$ times along every axis, then the reconstructed state $\tilde\rho_V$ approximates the true state $\rho$ to within a mean squared (Euclidean) distance $\epsilon_V^2/n$.
Furthermore, if $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

@@

To this end, we fix a particular set of measurement axes $V$, and consider performing $n$ measurements of spin projection along every axis $\v v\in V$, for a total of $N=\abs{V}\times n$ measurements.
Such a procedure is equivalent to making $N$ local measurements of the $N$-fold product state $\rho^{\otimes N}$.
For convenience, we index the tensor factors of $\rho^{\otimes N}$ by the integers $\p{i,j}$, with $i\in\set{1,2,\cdots,\abs{V}}$ specifying a measurement axis $\v v_i\in V$, and $j\in\set{1,2,\cdots,n}$ specifying the copy of $\rho$ prepared for the $j$-th measurement spin projection along a particular axis.
We then define the projectors $\Pi_{i\mu}\equiv\op{\mu_{\v v_i}}$ onto states $\ket{\mu_{\v v_i}}$ with definite spin projection $\mu$ along axis $\v v_i\in V$, and define $\Pi_{i\mu}^j$ to be an $N$-qudit operator with $\Pi_{i\mu}$ on the $\p{i,j}$-th tensor factor and the identity elsewhere.
We denote the experimental outcome of measuring $\Pi_{i\mu}$ in the $\p{i,j}$-th copy of $\rho$ by $\tilde \Pi_{i\mu}^j\in\set{0,1}$.
In other words, $\tilde \Pi_{i\mu}^j$ is the ``single-shot estimate'' of $\Pi_{i\mu}$, with $\tilde \Pi_{i\mu}^j=1$ if outcome $\mu$ was observed on the $\p{i,j}$-th experimental trial, and $\tilde \Pi_{i\mu}^j=0$ otherwise.
An empirical estimate of the expectation value $\bk{\Pi_{i\mu}}_\rho$ is provided by the fraction of times that outcome $\mu$ was observed when measuring spin projection along axis $\v v_i$, that is
\begin{align}
  \tilde \Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \tilde \Pi_{i\mu}^j
  \approx \f1n \sum_{j=1}^n \tr\p{\Pi_{i\mu}^j \rho^{\otimes N}}
  = \tr\p{\Pi_{i\mu} \rho}.
\end{align}
For reasons that will become clear shortly, it will be useful to think of $\tilde \Pi_{i\mu}$ as an empirical estimate of $\bk{\bar \Pi_{i\mu}}_{\rho^{\otimes N}}$, where
\begin{align}
  \bar \Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \Pi_{i\mu}^j
  \label{eq:mean_proj}
\end{align}
is the average of $\Pi_{i\mu}$ applied to all copies of $\rho$ for which spin projection is measured along the axis $\v v_i$.

@@

Finite sampling error (i.e.~shot noise) generally induces statistical error $\epsilon_\O$ into the empirical estimate $\tilde\O$ of an observable $\O$:
\begin{align}
  \epsilon_\O \equiv \tilde\O - \bk{\O},
\end{align}
where the single brackets $\bk{\cdot}$ denote an expectation value with respect to the measured quantum state.
On average, this statistical error will be zero, which is to say that
\begin{align}
  \bbk{\epsilon_\O} = \Bbk{\tilde\O - \bk{\O}}
  = \bk{\O - \bk{\O}}
  = 0,
\end{align}
where the double brackets $\bbk{\cdot}$ to denote statistical averaging over experimental trials.
However, the covariance between statistical errors $\epsilon_\O$ and $\epsilon_\Q$ on the empirical estimates $\tilde\O$ and $\tilde\Q$ of observables $\O$ and $\Q$ is
\begin{align}
  \bbk{\epsilon_\O \epsilon_\Q}
  = \Bbk{\p{\tilde\O - \bk{\O}} \p{\tilde\Q - \bk{\Q}}}
  = \bk{\p{\O - \bk{\O}} \p{\Q - \bk{\Q}}}
  = \bk{\O\Q} - \bk{\O} \bk{\Q}.
\end{align}
In the context of spin qudit tomography, we can therefore define the statistical error
\begin{align}
  \epsilon_{i\mu}
  \equiv \tilde \Pi_{i\mu} - \bk{\Pi_{i\mu}}_{\rho}
  = \tilde \Pi_{i\mu} - \bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}
\end{align}
in the empirical estimate of $\bk{\Pi_{i\mu}}_{\rho}$ or $\bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}$, and use Eq.~\eqref{eq:mean_proj} to expand
\begin{align}
  \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
  = \bk{\bar\Pi_{i\mu} \bar\Pi_{i'\mu'}}_{\rho^{\otimes N}}
  - \bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}
  \bk{\bar\Pi_{i'\mu'}}_{\rho^{\otimes N}}
  = \f1{n^2} \sum_{j,j'=1}^n
  \sp{\bk{\Pi_{i\mu}^j \Pi_{i'\mu'}^{j'}}_{\rho^{\otimes N}}
    - \bk{\Pi_{i\mu}^j}_{\rho^{\otimes N}}
    \bk{\Pi_{i'\mu'}^{j'}}_{\rho^{\otimes N}}}.
  \label{eq:proj_cov_start}
\end{align}
If $\p{i,j}\ne\p{i',j'}$, then $\Pi_{i\mu}^j$ and $\Pi_{i'\mu'}^{j'}$ address different tensor factors of the product state $\rho^{\otimes N}$, so the expectation value of their product factorizes due to the identity $\tr\sp{\p{A\otimes B}\p{A'\otimes B'}} = \tr\p{AA'}\times\tr\p{BB'}$.
This factorization can also be seen as a consequence of the fact that if $\p{i,j}\ne\p{i',j'}$, then $\Pi_{i\mu}^j$ and $\Pi_{i'\mu'}^{j'}$ are ``spatially separated'', which means that their expectation values with respect to the product state $\rho^{\otimes N}$ cannot have quantum correlations.
The terms in Eq.~\eqref{eq:proj_cov_start} with $\p{i,j}\ne\p{i',j'}$ therefore vanish, so
\begin{align}
  \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
  &= \delta_{ii'} \times \f1{n^2} \sum_{j=1}^n \sp{\bk{\Pi_{i\mu}^j \Pi_{i\mu'}^j}_{\rho^{\otimes N}}
    - \bk{\Pi_{i\mu}^j}_{\rho^{\otimes N}}
    \bk{\Pi_{i\mu'}^j}_{\rho^{\otimes N}}} \\
  &= \delta_{ii'} \times \f1n \sp{\bk{\Pi_{i\mu} \Pi_{i\mu'}}_\rho
    - \bk{\Pi_{i\mu}}_\rho \bk{\Pi_{i\mu'}}_\rho} \\
  &= \delta_{ii'} \times \f1n \sp{\delta_{\mu\mu'} p_\mu^i - p_\mu^i p_{\mu'}^i},
\end{align}
where $p_\mu^i \equiv \bk{\Pi_{i\mu}}_\rho = \bk{\mu_{\v v_i}|\rho|\mu_{\v v_i}}$ are the diagonal entries of $\rho$ in the basis of spin projections along axis $\v v_i$.

@@

Equipped with the bound on variances in Eq.~\eqref{eq:variance_bound}, we again consider the block-diagonalized form of the measurement matrix derived in Section \ref{sec:bound}:
\begin{align}
  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
  &&
  M_{V\ell} = \sum_{m,\v v} D^\ell_{m,0} \op{m}{\v v}.
\end{align}
The block diagonal structure of this matrix implies that, rather than an integer index $k\in\set{1,2,\cdots,d^2}$, we can identify the singular values and right singular vectors of $UM_V$ (and equivalently $M_V$) by a block index $\ell\in\set{0,1,\cdots,d-1}$ and a sub-index $m=\set{-\ell,-\ell+1,\cdots,\ell}$.
Making the replacement $k\to\p{\ell,m}$, we can thus repeat the derivations through Eq.~\eqref{eq:bound_eps} of Section \ref{sec:bound} without the assumption of Eq.~\eqref{eq:error_assumption}, finding that
\begin{align}
  \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
  = \sum_{i,i'} q_{\ell m i}^V \p{q_{\ell m i'}^V}^*
  \bbk{\epsilon_{i\ell} \epsilon_{i'\ell}}
  = \sum_i \abs{q_{\ell m i}^V}^2
  \bbk{\epsilon_{i\ell}^2}
  \le \f{2\ell+1}{nd} \times \gamma_\ell^2
  \times \p{1-\delta_{\ell,0}},
  \label{eq:bound_single}
\end{align}
where $i$ indexes an axis $\v v_i\in V$.
Altogether, we thus find that $n$ measurements of spin projection along every axis $\v v\in V$ allows us to construct an empirical estimate $\tilde\rho_V$ of an unknown state $\rho$ that is accurate to within a mean squared distance
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \sum_{\ell,m} \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
  < \f1{nd} \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2.
  \label{eq:bound_full}
\end{align}
We note that whereas the bound on $\bbk{\epsilon_{i\ell}^2}$ in Eq.~\eqref{eq:variance_bound} is tight, the bounds in Eq.~\eqref{eq:bound_single} and Eq.~\eqref{eq:bound_full} are not: the upper bound on $\bbk{\epsilon_{i\ell}^2}$ cannot be achieved simultaneously for all axes $\v v_i$ or degrees $\ell$.
In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound_full}.
